---
title: "Financial expertise and misreporting"
author: "Ian Gow"
date: "April 5, 2015"
output: pdf_document
---

## Research design

I conjecture that financial expertise of the board leads to less misreporting.
My research design is the *best*.
I regress a restatement indicator (*restate*) on the 
number of financial experts on the board (*num_exp*)
I include industry fixed effects using the Fama-French 48-industry classification (*ffic48*).

Here is the code to load the data.
```{r get_data}
library("foreign")
fin_exp <- read.dta("fin_exp.dta")
fin_exp$ffic48 <- as.factor(fin_exp$ffic48)
fin_exp$fyear <- as.factor(fin_exp$fyear)
```

To keep things simple, I don't transform my variables in any way.

The regression equation is:

$$\textit{restate}_{it} = \textit{num\_exp}_{it} \cdot \beta + \sum_{k \in \{1,\dots, 48\}} \textit{ind}_{ik} \cdot \gamma_k+ \epsilon_{it}$$

where $\textit{restate}_{it}$ is an indicator for whether firm $i$
restated financials related to fiscal year $t$, 
$\textit{num\_exp}_{it}$ is the number of financial experts on the board, 
and $\textit{ind}_{ik}$ is an indicator for firm $i$ being a member of industry $k$.

Here is the code for my regression:
```{r run_reg}
fm <- lm(restate ~ num_exp + factor(ffic48), data=fin_exp)
```

## Results

The results of my analysis are provided in Table 1 below (estimates industry fixed effects omitted for brevity). 
While the sign of the coefficient on \textit{num\_exp} is negative, 
I do not find a statistically significant association between financial expertise and restatements.
Perhaps my analysis is too simple.

```{r tabulate, message=FALSE, results='asis', echo=FALSE}
# install.packages("texreg")
library("texreg")
texreg(fm, omit.coef = "ffic48", label = "table:reg",
       caption = "Results of regressing \\emph{restate} on financial expertise")
```

```{r some_functions, include=FALSE}
# Simple regressions ----
getFitted <- function(list) {
  temp <- list()
  for (i in 1:length(list)) {
    temp[[i]] <- list[[i]][[1]]
  }
  return(temp)
}

# Get standard errors
getSEs <- function(a.list) {
  temp <- list()
  for (i in 1:length(a.list)) {
    temp[[i]] <- a.list[[i]][[2]][,2]
  }
  return(temp)
}

# Get p-values
getPs <- function(a.list) {
  temp <- list()
  for (i in 1:length(a.list)) {
    temp[[i]] <- a.list[[i]][[2]][,4]
  }
  return(temp)
}
```

## Matt's approach
```{r matt, warning=FALSE, message=FALSE}
fin_exp$aaer_or_hlm <- fin_exp$aaer | fin_exp$restate_hlm
source("http://iangow.me/code/cluster2.R")

lhs.vars <- c("restate", "restate_hlm", "aaer", "aaer_or_hlm")

fit.model <- function(lhs) {
    formula <- paste(lhs, "~ num_exp_aud + fyear + ffic48")
    fm <- glm(formula, data=fin_exp, family = binomial(link = "probit"))
    fm.se <- coeftest.cluster(fin_exp, fm, cluster1="firmid")
    return(list(fm, fm.se))
}

fms <- lapply(lhs.vars, fit.model)


```{r matt_out, results="asis", message=FALSE}
library("stargazer")
stargazer(getFitted(fms), omit=c("^fyear", "^ffic"), 
          title = "Matt's regressions",
          se = getSEs(fms), p = getPs(fms), header = FALSE,
          keep.stat=c("n", "adj.rsq"))


```